---
title: "EasySentiment Functions"
author: "Diego Ford"
date: "2024-03-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Overview

This Document will serve to explain and show the code behind the functions created for my EasySentiment package. **Please note that this package is in version 0.0.0.900 at the time of writing and things may change by the final release**



#Functions

##get_sentiments function

### Purpose

The purpose of this function is to get the sentiments of text inputted into it. It will quickly and easily return the sentiments of text inputted into it. 

### Inputs and Outputs

The function will take in a vector of strings and returns a tibble containing the sentiments for each word in each string. It will also take in as an input the sentiment dictionary you would like to use, which will be *"bing"* *"nrc"* *loughran* or *afinn*, support for other lexicons or custom lexicons may be added in the future.

### Function Construction

```{r get_sentiments function}
library(dplyr)

get_sentiments <- function(text, lexicon) {
  text <- text
  
  if(tolower(lexicon) %in% c("bing", "afinn", "nrc", "loughran")) {
    sentiments <- tidytext::get_sentiments(lexicon)
  } else {
    stop("An invalid sentiment lexicon was selected")
  }
  
  # Tokenize
  text <- data.frame(text) 
  text <- text %>%
    dplyr::mutate(indexes = dplyr::row_number())
  
  colnames(text) <- c("text", "indexes")
  text_tokens <- text %>%
    tidytext::unnest_tokens(output = word, input = text, token = "words")
  
  text_sentiments <- text_tokens %>%
    dplyr::inner_join(sentiments)
  
  return(text_sentiments)
}



```

### example

```{r get_sentiments example}

example_strings <- c("I am really angry right now", "I am very very happy", "Rage and mixed feelings", "i lied")
example_sentiments <- get_sentiments(example_strings, "bing")
head(example_sentiments)
```

## Sentiment_Vis Function

### Purpose

This function allows for several visuals regarding the distribution of sentiments in the corpora.

### Inputs and Outputs

The inputs for this function is *index* which is a unique index for each text document and *vis_type* which tells which type of vis you would like generated right now you can chose sentiment_hist or word_hist, other visuals will be added before final release the final input is *sentiments* which is the sentiments associated with each text document. The output is a graph.

### Function Construction
```{r sentiment_vis}
sentiment_vis <- function(sentiments, index, vis_type){
  data <- data.frame(index, sentiments)
  
  #get counts of sentiments
  if(vis_type == "sentiment_hist"){
    data <- data %>%
      group_by(index, sentiments) %>%
      count()
  
  ggplot2::ggplot(data, ggplot2::aes(x = as.factor(index), y = n)) +
    ggplot2::geom_col(ggplot2::aes(fill = as.factor(sentiments)), position = "dodge") +
    ggplot2::labs(xlab = "indexes", ylab = "sentiments", fill = "sentiment")
  }
}
```

### example

```{r example}
plot <- sentiment_vis(example_sentiments$sentiment, example_sentiments$indexes, vis_type = "sentiment_hist")
plot
```





